---
title: "exam_2_answer"
author: "Søren Ahlers"
format: 
  html:
    embed-resources: true
editor: visual
---

## Quarto

```{r}
dat <- doBy::personality
set.seed(101) # for reproducibility
i_train <- sample(nrow(dat), .5*nrow(dat))
train <- dat[i_train,]
test <- dat[-i_train,]
```

## Exercise 1

```{r}
library(ggplot2)
doBy::response_plot(
  dat,
  easygon ~ .,
  geoms = ggplot2::geom_point(),
  #global_aes = ggplot2::aes()
  ncol = 5)#+
  

```

## Exercise 2

Missing

## Exercise 3

```{r}
 tree1 <- rpart::rpart(data = dat, formula = easygon ~ .)
library(rpart.plot)
rpart.plot(tree1)
```

# Part II

## Exercise 1

## Exercise 1

Det lægges til grund, at blodtryk, hudtykkelse, bmi, insulin og glukose alle skal være skarpt større en 0, da dette er en forudsætning for at patienten er i live.

```{r}
# Absolut sti til data, uanset system:
sti_dia <- here::here("diabetes.csv")

dat_raw <- readr::read_csv(
  file = sti_dia)



cols <- c("BloodPressure", "SkinThickness", "BMI", "Insulin", "Glucose")

# Værdier erstattes med NA:
dat <- dat_raw |> 
  # For at være i live skal man have hud og blodtryk, dvs. hvis disse er 0, så
  # det en anomali:
  dplyr::mutate(
    dplyr::across(dplyr::all_of(cols), ~ dplyr::na_if(., 0))
  )

dat <- dat |> 
  # Der filtreres, så alle rækker med NA-værdier fjernes:
  dplyr::filter(!dplyr::if_any(dplyr::all_of(colnames(dat)), ~ is.na(.)))


dim(dat)

```

## Exercise 2

```{r}
library(ggplot2)
library(patchwork)

geom_diabetes <- function(x_var) {
  p1 <- dat |> 
  ggplot2::ggplot(aes(y = Outcome, x = .data[[x_var]]))+
  #ggplot2::geom_point() +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.2)+
  geom_smooth(method = "loess", formula = "y ~ x")
  
  p1
}

plotvar <- colnames(dat)[1:8]

hest <- lapply(plotvar, geom_diabetes)


patchwork::wrap_plots(hest, ncol = 3)
```

## Exercise 3

Ensure `Outcome` is a `factor`:

```{r}
library(dplyr)
dat <- dat |> 
  mutate(Outcome = as.factor(Outcome))
```

and split the data in a training and test set as follows:

```{r}
set.seed(202) # for reproducibility
i_train <- sample(nrow(dat), .5*nrow(dat))
train <- dat[i_train,]
test <- dat[-i_train,]
```

Specify a number of classification models and report the corresponding confusion matrix for the test data. You must include at least one of each of the following:

1.  logistic regression with stepwise selection (glm / step)

2.   regression tree (rpart)

3.  random forest (randomForest)

4.  linear/quadratic/regularized discriminant analysis (use sda_formula() for regularized – see updated notes).

### 3.2 Regression tree
